Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-01-09 Manuscript/manuscript.Rmd
edits

## SMOTE re-sampling, distance-based associations of TAD boundaries and genomic annotations, and 10 kb resolutions of Hi-C data provided optimal model performance

@@

Interestingly, we found that performances peaked at 10 kb resolution and then degraded as data resolution decreased. The reason for the improved performance between 5 kb and 10 kb data resolutions can be due to the sparsity in the interaction frequency matrices at finer resolutions. [??? Quantify sparsity] This sparsity may lead to inaccurate TAD calling and, subsequently, to worse model performance. These results suggest that finer resolution of Hi-C data is not always optimal for accurate TAD calling and prediction, and sparsity of the data needs to be considered. Additionally, performances tended to be much more variable between chromosomes at lower resolutions, evidenced by the wider error bars for the 50 kb and 100 kb resolutions. In summary, our results imply that using random under- and over-sampling, distance-type predictors, and the data at 10kb resolution results in an optimally performing cell line-specific random forest-based TAD boundary prediction model.

<!--
![**Models built using distance-type predictors combined with SMOTE re-sampling yielded optimal MCCs across resolution.** Barplots comparing the performances of random forest classifiers built using different combinations of data level characteristics. Within each re-sampling technique (A-D), performances are compared for each type of feature space: OC (red), OP (green), and Distance (blue), across resolutions. Performances were evaluated using MCCs and aggregated by averaging across each chromosome, with error bars representing 1 standard deviation away from the mean.](Figures/model_performances_v3_bars.png)
-->

>>>>

## Random under-sampling, distance-based predictors, and 10 kb resolutions of Hi-C data provided the optimal data level characteristics for TAD boundary prediction

@@

Interestingly, we found that performances peaked at 10 kb resolution and then degraded as data resolution decreased. The reason for the improved performance between 5 kb and 10 kb data resolutions can be due to the sparsity in the interaction frequency matrices at finer resolutions. This sparsity may lead to inaccurate TAD calling and, subsequently, to worse model performance. These results suggest that finer resolution of Hi-C data is not always optimal for accurate TAD calling and prediction, and sparsity of the data needs to be considered. Additionally, performances tended to be much more variable between chromosomes at lower resolutions, evidenced by the wider error bars for the 50 kb and 100 kb resolutions. In summary, our results imply that using random under- and over-sampling, distance-type predictors, and the data at 10kb resolution results in an optimally performing cell line-specific random forest-based TAD boundary prediction model.

![**Models built using distance-type predictors combined with random under-sampling yielded optimal predictive performance.** Line plots comparing the performances of random forest classifiers built using different combinations of data level characteristics. Performances are compared across resolution, within each predictor-type: (A) OC, (B) OP, and (C) Distance, and between each re-sampling technique: no re-sampling (None; red), random over-sampling (ROS; green), random under-sampling (RUS; blue), and synthetic minority over-sampling (SMOTE; purple). Performances were evaluated using MCCs and aggregated by averaging across each chromosome, with error bars representing 1 standard deviation away from the mean.](Figures/performances.png)
